---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r }
library(Hmisc)
library(dplyr)
library(ggplot2)
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip", mode="wb")
# unzip data and read 
unzip("activity.zip")
data <- read.csv("activity.csv", header = TRUE)

str(data)
head(data)
data$date <- as.Date(data$date)
data1 <- na.omit(data)
steps <- aggregate(data1$steps, by = list(Date = data1$date), FUN = sum)
colnames(steps) <- c("day", "steps")
```

## What is mean total number of steps taken per day?

```{r}
mean(steps$steps)
median(steps$steps)
```

#####Histogram of the total number of steps taken each day
```{r}
hist(steps$steps, xlab ="total steps each day", main ="Histogram of total 
     steps each day")
```


```

## What is the average daily activity pattern?
```{r}
interval <- aggregate(data1$steps, by = list(data1$interval), FUN =mean)
colnames(interval ) <- c("interval", "steps")
```
##### Time series plot of 5 min intervals
```{r}
plot(interval$interval, interval$steps, type ="l",
         col ="blue", xlab ="intervals" , ylab ="avg. steps")
     
interval[which(interval$steps== max(interval$steps)),]

```


## Imputing missing values
```{r}
sum(is.na(data))
imputed_data <- data
imputed_data$steps<- impute(data$steps, fun=mean)
imputed_steps <- aggregate(imputed_data$steps, by = list(Date = imputed_data$date), FUN = sum)
colnames(imputed_steps) <- c("day", "steps")
```
#####  Histogram of the total number of steps taken each day in imputed data 
```{r}
hist(imputed_steps$steps, xlab ="total steps each day", main ="Histogram of total 
     steps each day")
```
##### MEan and meadian of steps
```{r}
mean(imputed_steps$steps)
median(imputed_steps$steps)

```
## Are there differences in activity patterns between weekdays and weekends?
```{r}
imputed_data$day <- weekdays(imputed_data$date)
imputed_data$day<- ifelse(imputed_data$day=="Saturday" | imputed_data$day=="Sunday", "Weekend", "Weekday" )



imputed_data_day <- aggregate(imputed_data$steps , by= list(imputed_data$day, imputed_data$interval), FUN = mean)
names(imputed_data_day) <- c("weekend", "interval", "steps")
```
##### 2. Panel plot containing a time series plot
```{r}
ggplot(imputed_data_day, aes(x=interval, y=steps, color=weekend)) + geom_line()+
        facet_grid(weekend ~.) + xlab("Interval") + ylab(" Avg .Steps") +
        ggtitle("Comparison of Activity between Weekend and Weekdays")
```


